---
title: "AE 11: Multinomial classification"
---

::: callout-important
Go to the [course GitHub organization](https://github.com/STA210-Summer22) and locate the repo titled `ae-11-volcanoes-YOUR_GITHUB_USERNAME` to get started.
:::

## Packages

```{r}
#| label: load-pkgs
#| warning: false
 
library(tidyverse)
library(tidymodels)
library(knitr)
library(colorblindr)
```

## Data

For this application exercise we will work with a dataset of on volcanoes. The data come from [The Smithsonian Institution](https://volcano.si.edu/) via [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12/readme.md).

```{r}
volcano <- read_csv(here::here("ae", "data/volcano.csv"))
```

First, a bit of data prep:

```{r}
volcano <- volcano %>%
  mutate(
    volcano_type = case_when(
      str_detect(primary_volcano_type, "Stratovolcano") ~ "Stratovolcano",
      str_detect(primary_volcano_type, "Shield") ~ "Shield",
      TRUE ~ "Other"
    ),
    volcano_type = fct_relevel(volcano_type, "Stratovolcano", "Shield", "Other")
  ) %>%
  select(
    volcano_type, latitude, longitude, 
    elevation, tectonic_settings, major_rock_1
    ) %>%
  mutate(across(where(is.character), as_factor))
```

## Exploratory data analysis

1.  Create a map of volcanoes that is faceted by `volcano_type`. First, we create a map here. 

```{r}
world <- map_data("world")

world_map <- ggplot() +
  geom_polygon(
    data = world, 
    aes(
      x = long, y = lat, group = group),
      color = "white", fill = "gray50", 
      size = 0.05, alpha = 0.2
    ) +
  theme_minimal() +
  coord_quickmap() +
  labs(x = NULL, y = NULL)
world_map
```

Then you need to add a layer to the map using `geom_point`, and reproduce the following plot.

```{r}
#| echo: false
world_map +
  geom_point(
    data = volcano,
    aes(x = longitude, y = latitude,
        color = volcano_type, 
        shape = volcano_type),
    alpha = 0.5
  ) +
  facet_wrap(~volcano_type) +
  scale_color_OkabeIto()
```

## Build a new model

2.  Build a new model that uses a recipe that includes geographic information (latitude and longitude). How does this model compare to the original one presented on slides? **Note:**\
    Use the same test/train split as well as same cross validation folds. Use seed 1234 for data splitting, and 9876 for cv.

```{r}
```

New recipe, including geographic information:

```{r}
```

Combine the original model specification as the one on slides and new recipe to obtain new workflow:

```{r}

```

Fit resamples, and save it to `volcano\_fit\_rs2`

```{r}

```

Collect metrics:

```{r}

```

Draw ROC curves:

```{r}

```

Compare these with the ones on slides. Comments on it. Which model do you prefer?

## Acknowledgement

This exercise was inspired by [https://juliasilge.com/blog/multinomial-volcano-eruptions](https://juliasilge.com/blog/multinomial-volcano-eruptions/).
